Semantic integration of physiology phenotypes with an application to the Cellular Phenotype Ontology

### 2012-04-02 paper/drinking-bioinformatics.tex

Amongst abnormal physiology, we distinguish between abnormalities in a
{\em single occurrence} of a cell component's functioning and an
abnormal {\em pattern of multiple occurrences} of a cell component's
functioning. For example, abnormalities in cell division resulting in
{\em Aneuploidy} refer to abnormalities of {\em Cell division}
processes, while an {\em Increased rate of cell division} refers to an
abnormality in the pattern of occurrence of multiple cell division
processes. In the CPO we follow the GO and represent abnormalities in
the pattern of occurrence of $X$ as abnormalities of {\em Regulation
  of $X$} processes.  In particular, an {\em Increased rate of cell
  division} is not an attribute of {\em Cell division} processes, but
rather arises from the collection of all {\em Cell division} processes
that occur within an organism at a given time. In the
GO, {\em Regulation of $X$} processes refer to those processes that
determine how often and the way in which one or more $X$ processes
occur. Therefore, we assign the attribute of {\em Increased rate of
  cell division} to {\em Regulation of cell division} processes.

@@

\section{Acknowledgement}
Funding for RH is provided by the European Commission's 7th Framework
Programme, RICORDO project, grant number 248502. MAH is supported by
Wellcome Trust, grant WT090548MA. HH is supported by the Institute for
Medical Informatics, Statistics and Epidemiology, University of
Leipzig.  Funding for GR is provided by the European Union's Seventh
Framework Programme (FP7/2007-2013) under grant agreement number
258068, EU-FP7-Systems Microscopy NoE.

>>>>

Amongst abnormal physiology, we distinguish between abnormalities in a
{\em single occurrence} of a cell component's functioning and an
abnormal {\em pattern of multiple occurrences} of a cell component's
functioning. For example, abnormalities in cell division resulting in
{\em Aneuploidy} refer to abnormalities of {\em Cell division}
processes, while an {\em Increased rate of cell division} refers to an
abnormality in the pattern of occurrence of multiple cell division
processes. In the CPO we follow the GO and represent abnormalities in
the pattern of occurrence of $X$ as abnormalities of {\em Regulation
  of $X$} processes.  In particular, an {\em Increased rate of cell
  division} is not an attribute of {\em Cell division} processes, but
rather arises from the collection of all {\em Cell division} processes
that occur within an organism at a given time. In the GO, {\em
  Regulation of $X$} processes refer to those processes that determine
how often and in which way one or more $X$ processes occur. Therefore,
we assign the attribute of {\em Increased rate of cell division} to
{\em Regulation of cell division} processes.

@@

\section{Acknowledgement}
We thank Michel Dumontier for valuable discussions about the formal
representation of the axioms.

Funding for RH is provided by the European Commission's 7th Framework
Programme, RICORDO project, grant number 248502. MAH is supported by
Wellcome Trust, grant WT090548MA. HH is supported by the Institute for
Medical Informatics, Statistics and Epidemiology, University of
Leipzig.  Funding for GR is provided by the European Union's Seventh
Framework Programme (FP7/2007-2013) under grant agreement number
258068, EU-FP7-Systems Microscopy NoE.
